---
title: "Daten"
author: "Justin Lampman"
date: "2025-02-26"
output: html_document
---

For the code to work without loading any data previously, you first have to run PCA.Rmd, then you have to run Daten.Rmd to have all data. To run the models you have to run rf.Rmd. For the plots you need to run plots.Rmd. 
To load all data without running the code, you can use the csv files in the csv data folder and run them in the second chunk. The models are saved in the models folder. The plots are saved in the plots folder. The data is saved in the data folder.

packages
```{r}
library(worldfootballR)
library(tidyverse)
library(ggrepel)
library(stringi)
library(stringr)
library(worldfootballR)
library(corrr)
library(ggcorrplot)
library(factoextra)
library(purrr)
library(vivid)
library(scales)
library(reshape2)  # For data transformation
library(viridis) 
library(ranger)
library(Metrics)
mapping = read.csv("csv data/names.csv")# Mapping for clubs of transfermakt data and fbref data
```

If you do not want to run the code, you can also use the csv files in the csv data folder and load these to have all the data, to run the models and get the plots
```{r}
pca_join_age_diff2total <- read.csv("csv data/pca.csv")
dataa <- read_csv("csv data/dataa123.csv")%>% select(-1)
#
rfreg_tot_vivi2 <- read.csv("csv data/rfreg_tot_vivi2.csv")
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
data_butlook <- read_csv("csv data/data.csv")
#
mid_data2reg <- read.csv("csv data/mid_data2reg.csv")
df_data2reg <- read.csv("csv data/df_data2reg.csv")
tot_data2reg <- read.csv("csv data/vtot_data2reg.csv")
fw_data2reg <- read.csv("csv data/fw_data2reg.csv")
fw_data2bin <- read.csv("csv data/fw_data2bin.csv")
mid_data2bin <- read.csv("csv data/mid_data2bin.csv")
df_data2bin <- read.csv("csv data/df_data2bin.csv")
tot_data2bin <- read.csv("csv data/tot_data2bin.csv")
#
combined_varimp_bin2 <- read.csv("csv data/combined_varimp_bin2.csv")
combined_varimp_reg2 <- read.csv("csv data/combined_varimp_reg2.csv")
#
regbin <- read.csv("csv data/regbin.csv")
#
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
train_data_df2 <- read.csv("csv data/train_data_df2.csv")
test_data_df2 <- read.csv("csv data/test_data_df2.csv")
test_data_fw2 <- read.csv("csv data/test_data_fw2.csv") 
train_data_fw2 <- read.csv("csv data/train_data_fw2.csv")
train_data_mid2 <- read.csv("csv data/train_data_mid2.csv")
test_data_mid2 <- read.csv("csv data/test_data_mid2.csv")
train_data_tot2 <- read.csv("csv data/train_data_tot2.csv")
test_data_tot2 <- read.csv("csv data/test_data_tot2.csv")
#
df_data <- read.csv("csv data/df_data.csv")
mid_data <- read.csv("csv data/mid_data.csv")
fw_data <- read.csv("csv data/fw_data.csv")
tot_data <- read.csv("csv data/tot_data.csv")
#
rfbin_fw2 <- readRDS("models/rfbin_fw2.rds")
rfbin_df2 <- readRDS("models/rfbin_df2.rds")
rfbin_mid2 <- readRDS("models/rfbin_mid2.rds")
rfbin_tot2 <- readRDS("models/rfbin_tot2.rds")
rfreg_fw2 <- readRDS("models/rfreg_fw2.rds")
rfreg_df2 <- readRDS("models/rfreg_df2.rds")
rfreg_mid2 <- readRDS("models/rfreg_mid2.rds")
rfreg_tot2 <- readRDS("models/rfreg_tot2.rds")
# load these for the data from Transfermarkt and Fbref
all_val <- read.csv("csv data/all_val.csv")
final_combined_data <- read.csv("csv data/final_combined_data.csv")
joined_data = read.csv("csv data/team_data_pca.csv")
```


get data from fbref
```{r}
data_standard <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "standard", team_or_player= "player")

data_shooting<- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "shooting", team_or_player= "player")

data_passing <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "passing", team_or_player= "player")

data_passing_types <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "passing_types", team_or_player= "player")

data_gca <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "gca", team_or_player= "player")

data_defense <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "defense", team_or_player= "player")

data_possession <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "possession", team_or_player= "player")

data_playing_time <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "playing_time", team_or_player= "player")

data_misc <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "misc", team_or_player= "player")

#data_keepers <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "keepers", team_or_player= "player")

#data_keepers_adv <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= "keepers_adv", team_or_player= "player")
#keepers are excluded due to the fact that they are not relevant for the analysis
```

combine all datasets 
```{r}
datasets <- list(data_defense, data_gca, data_misc, data_passing, data_passing_types, 
                  data_possession, data_shooting, data_standard)

# Keep only the first 8 columns of each dataset
first_8_cols <- lapply(datasets, function(df) df[, 1:8])

base_cols <- first_8_cols[[1]]

# Keep all remaining columns from each dataset after the first 8
extra_cols <- lapply(datasets, function(df) df[, -(1:8)])

combined_extra_cols <- do.call(cbind, extra_cols)

unique_extra_cols <- combined_extra_cols[, !duplicated(colnames(combined_extra_cols))]

# Combine the base columns with the unique extra columns
final_combined_data <- cbind(base_cols, unique_extra_cols)

final_combined_data = left_join(final_combined_data,data_playing_time, by = c("Player","Season_End_Year","Comp","Squad", "Url","Nation","Pos","Age", "Born"))

#write.csv(final_combined_data, "csv data/final_combined_data.csv", row.names = FALSE)
```

Get transfervalues for each year
```{r}
all_val17 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2017)

all_val18 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2018)

all_val19 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2019)

all_val20 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2020)

all_val21 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2021)

all_val22 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2022)

all_val23 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2023)

all_val24 <- tm_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),start_year = 2024)

all_val <- rbind(all_val17,all_val18, all_val19, all_val20, all_val21, all_val22, all_val23, all_val24)
all_val2 <- rbind(all_val17,all_val18, all_val19, all_val20, all_val21, all_val22, all_val23, all_val24)
#write.csv(all_val, "csv data/all_val.csv", row.names = FALSE)
```

make the data from transferamrkt nice and formatable 
```{r}
all_val2 = all_val
# Define regex pattern for detecting numbers
market_value_pattern <- "[0-9]"

for (i in 1:(nrow(all_val) - 1)) {
  if (grepl(market_value_pattern, all_val$player_name[i])) {
    all_val$MarketValue[i - 1] <- all_val$player_name[i]
    all_val$player_name[i] <- NA
  } 
  else if (!grepl(market_value_pattern, all_val$player_name[i]) && 
           !grepl(market_value_pattern, all_val$player_name[i + 1])) {
    # Assign "0€" if both the current and the next row do not contain numbers
    all_val$MarketValue[i] <- "0€"
  }
}

# Remove rows where 'player_name' is (not) NA
all_val <- all_val[!is.na(all_val$player_name), ]
all_val$rownr = seq.int(nrow(all_val))

all_val = right_join(all_val,mapping, by = "squad") %>%
dplyr::select(-squad) %>%
 rename(squad = Squad)

all_val2 = right_join(all_val2,mapping, by = "squad") %>%
dplyr::select(-squad) %>%
 rename(squad = Squad)

```


put together with data both
```{r}
normalize_name <- function(name) {
  return(stri_trans_general(name, "Latin-ASCII"))
}

all_val = all_val %>% 
  dplyr::mutate(comp_name = gsub("LaLiga", "La Liga", comp_name),
         season_start_year = season_start_year + 1,
         player_name = sapply(player_name, normalize_name),
         key = paste(squad,player_name,season_start_year,comp_name,sep = "_"))%>% 
  arrange(squad, season_start_year) %>% 
  select(-player_position)

############################################
#make that player position is right and in there
all_val_summary <- all_val %>%
  group_by(squad, season_start_year) %>%
  summarise(count = n() , .groups = "drop")


all_val2_first_position <- all_val2 %>%
    mutate(season_start_year = season_start_year + 1)%>% #shift the year
  inner_join(all_val_summary, by = c("squad", "season_start_year")) %>% 
  group_by(squad, season_start_year) %>%
  group_modify(~ head(.x, .x$count.y[1])) %>%   # Subset each group based on its own count value
  ungroup() %>%
  select(squad, season_start_year, player_position)%>% 
  arrange(squad, season_start_year)

all_val = cbind(all_val,all_val2_first_position[3])
############################################

final_combined_data = final_combined_data %>%
  dplyr::mutate(Player = sapply(Player, normalize_name),
         key = paste(Squad,Player,Season_End_Year,Comp,sep = "_"))

whole_data = left_join(final_combined_data,all_val,by = "key")

```

make data not that players arent duplicate
```{r}
data = whole_data[!is.na(whole_data$MarketValue),]
data = data[c(1:180,197:203)]

data = data[!duplicated(data[1:182]),]#without last coloumn

data2 = data #here the duplicates are now lost

data2 = data %>% select(Player,Comp,Squad,Age,MarketValue,player_url,Pos,rownr)

#make transfer valuen numeric
data2$MarketValue <- sapply(data2$MarketValue, function(x) {
  if (grepl("k", x, ignore.case = TRUE)) { 
    # For values with "k"
    as.numeric(gsub("[^0-9.]", "", x)) * 1e3 
  } else if (grepl("M", x, ignore.case = TRUE)) { 
    # For values with "M"
    as.numeric(gsub("[^0-9.]", "", x)) * 1e6 
  } else { 
    # Default case (numeric values without a suffix)
    as.numeric(gsub("[^0-9.]", "", x)) 
  }
})


ggplot((data2), aes(x = Age, y = MarketValue, group = interaction(Player, Squad), color = Player)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::dollar_format(scale = 1, suffix = "€")) +
  labs(
    title = "Market Value of Players by Age",
    x = "Age",
    y = "Market Value (in €)",
    color = "Player"
  ) +
  theme_minimal()+
  theme(legend.position = "none")
#based on this avererage gain or loss of a player value by age as coefficient of a linear regression model
#how to transfer data that you have a single player and then if he gained or lost value, transfrom the data to reflect that
#summarize paper quick and write about the transfer value and the age of the player and how to model how good a player is based on transfer value

#write.csv(data, "csv data/data.csv", row.names = FALSE)
```

avergae loss by age 
```{r}
#Filter and take the first entry for each player and age
datalm <- data2 %>%
  group_by(player_url, Age) %>%
  summarise(MarketValue = first(MarketValue), .groups = "drop")

#Pivot data to make each player a row, with columns for ages
pivoted_data <- datalm %>%
  pivot_wider(
    names_from = Age,  # Make columns based on ages
    values_from = MarketValue,  # Fill columns with market values
    names_prefix = "Age_"
  )

#Reorder columns based on age numerically
pivoted_data <- pivoted_data %>%
  select(player_url, order(as.numeric(gsub("Age_", "", names(.)))))

#Calculate percentage change for each player
long_data <- datalm %>%
  arrange(player_url, Age) %>%
  group_by(player_url) %>%
  mutate(
    pct_change = (MarketValue / lag(MarketValue) - 1) * 100  # Percentage change
  ) %>%
  ungroup()

#Plot each player's data
ggplot(long_data, aes(x = Age, y = MarketValue, color = player_url, group = player_url)) +
  geom_line() +  # Connect points for each player
  geom_point() +  # Show individual points
  labs(
    title = "Player Market Value Over Time",
    x = "Age",
    y = "Market Value",
    color = "Player"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


#Filter out NA and Inf in percentage change and consider only losses
filtered_data2 <- long_data %>%
  filter(!is.na(pct_change) & !is.infinite(pct_change) )

avg_loss_by_age2 <- filtered_data2 %>% 
  group_by(Age) %>% 
  summarise(
    avg_pct_loss = median(1 + pct_change / 100, na.rm = TRUE),
    .groups = "drop"
  )
```

make the data like i want it with zzzinital value
```{r}
datause <- data
datause = datause %>% distinct()
datause$Season_End_Year <- as.numeric(datause$Season_End_Year)
# Fill missing years for each player_url
data_filled <- datause %>% 
  arrange(player_url, Season_End_Year) %>% # Sort by player_url and year
  group_by(player_url) %>%
  complete(Season_End_Year = 2018:2024) %>% # Ensure years 2018 to 2024 exist
  mutate(
    Squad = ifelse(is.na(Squad), "no info", Squad), # Set Squad to "no info" for new rows
    Comp = ifelse(is.na(Comp), coalesce(lag(Comp), lead(Comp)), Comp),
    MValue = ifelse(is.na(MarketValue), coalesce(lag(MarketValue), lead(MarketValue)), MarketValue),
   ) %>%
  ungroup()%>%
mutate(rowid = row_number())

transitions <- data_filled %>%
  arrange(player_url, Season_End_Year) %>%
  group_by(player_url) %>%
  mutate(
    Previous_Squad = lag(Squad), # Look at the previous Squad
    Is_Change = (!is.na(Previous_Squad) & Squad != Previous_Squad & Squad != "no info")
  ) %>%
  filter(Is_Change) %>%
  select(-Is_Change,-Previous_Squad) %>%
  ungroup() #transitions is the data after the transfer happend

rows = (transitions$rowid -1)

transitions2 <- data_filled %>% filter(rowid %in% rows) %>% select(-MValue,- rowid)
transitions2$Squad <- "zzzinitial value"
data_filled <- data_filled[, colnames(datause)]

data_try = rbind(transitions2,datause) %>% arrange(player_url, Season_End_Year,rownr,Squad)

```

make the data so that you can use it
```{r}
rows_to_remove <- data_try %>%
  filter(Squad == "zzzinitial value") %>%
  group_by(Squad, Season_End_Year, player_url) %>%
  filter(n() > 1) %>%  # Only keep groups with more than one row
  arrange(Squad, Season_End_Year, player_url, rownr) %>%  # Sort by rownr (ascending order)
  slice_tail(n = 1) %>%  # Take the last (highest rownr) for each group
  select(rownr)  # Only keep the rownr to remove

# Remove rows with higher rownr
data_try <- data_try %>%
  filter(!(Squad == "zzzinitial value" & rownr %in% rows_to_remove$rownr))


# Add global and local transfer classification
data_try <- data_try %>%
  arrange(player_url, Season_End_Year,rownr,Squad) %>%  # Sort by player_url and Season_End_Year
  group_by(player_url) %>%
  mutate(
    Transfer_Group = cumsum(Squad == "zzzinitial value"),  # Group transfers by "zzzinitial value"
    Is_Transfer = ifelse(Squad == "zzzinitial value" | lag(Squad, default = Squad[1]) == Squad, "Same Transfer", "New Transfer")
  ) %>%
  ungroup() %>%
  mutate(
    global_transfer_index = dense_rank(paste(player_url, Transfer_Group))  # Global transfer index
  )
indices_to_remove <- data_try %>%
  filter(is.na(MarketValue)) %>%
  select(global_transfer_index)

#Remove rows with these Global_Transfer_Index values
data_try_cleaned <- data_try %>%
  filter(!global_transfer_index %in% indices_to_remove$global_transfer_index) %>% mutate(
    global_transfer_index = dense_rank(paste(player_url, Transfer_Group))  # Global transfer index
  )

data_try_cleaned$Age = as.numeric(data_try_cleaned$Age)

length(unique(data_try_cleaned$global_transfer_index))

data_try_cleaned$MarketValue <- sapply(data_try_cleaned$MarketValue, function(x) {
  if (grepl("k", x, ignore.case = TRUE)) { 
    # For values with "k"
    as.numeric(gsub("[^0-9.]", "", x)) * 1e3 
  } else if (grepl("M", x, ignore.case = TRUE)) { 
    # For values with "M"
    as.numeric(gsub("[^0-9.]", "", x)) * 1e6 
  } else { 
    # Default case (numeric values without a suffix)
    as.numeric(gsub("[^0-9.]", "", x)) 
  }
})

filtered_data01 <- data_try_cleaned %>%
  group_by(global_transfer_index) %>%
  filter(any(Squad == "zzzinitial value")) %>% #Keep only groups where `Squad` has "zzzinitial value"
  ungroup() %>% arrange(player_url, Season_End_Year,rownr,Squad)
 
filtered_data_cleaned <- filtered_data01 %>%
  filter(!is.na(Age))

data_transfers = filtered_data_cleaned
```

calculate the transfer value of the player
```{r}
calculate_transfer <- function(data_transfers, avg_loss_by_age) {
  results <- data.frame(global_transfer_index = integer(), Transfer_Value = numeric(), stringsAsFactors = FALSE)
  
  # Loop over each unique global_transfer_index in the dataset
  for (transfer_index in unique(data_transfers$global_transfer_index)) {
    transfer_data <- data_transfers[data_transfers$global_transfer_index == transfer_index, ]
    
    n_transfers <- nrow(transfer_data)  
    if (n_transfers < 2) next  # Skip if fewer than 2 transfers
    
    # Calculate the discount factor dynamically
    discount_factor <- sum(1 / 1:n_transfers) - 1
    
    # Extract values for the first transfer
    mv1 <- transfer_data$MarketValue[[1]]
    if (mv1 == 0) mv1 <- 10000  # Default value if MarketValue is 0
    
    age_idx1 <- transfer_data$Age[2] - 15  # Index for first transfer
    if (age_idx1 <= 0 || age_idx1 > nrow(avg_loss_by_age)) next  # Skip if invalid age
    
    # Start calculating y with the first value
    y <- -(mv1 * avg_loss_by_age$avg_pct_loss[age_idx1])
    
    # Loop over the remaining years after transfers
    for (j in 2:n_transfers) {
      mv <- transfer_data$MarketValue[j]
      age_idx <- transfer_data$Age[j] - 15  # Adjust the index
      
      if (age_idx > 0 && age_idx <= nrow(avg_loss_by_age)) {  # Ensure valid age
        weight <- (1 / j) / discount_factor  # Calculate weight
        y <- y + mv * avg_loss_by_age$avg_pct_loss[age_idx] * weight
      }
    }
    results <- rbind(results, data.frame(global_transfer_index = transfer_index, Transfer_Value = y))
  }
  
  return(results)
}


results = calculate_transfer(data_transfers, avg_loss_by_age2)
```

```{r}
results <- right_join(results,data_transfers, by = "global_transfer_index")

#add extra comp where he was before
results$Comp_going <- NA
# Identify rows where Squad is "zzzinitial value"
zzz_rows <- which(results$Squad == "zzzinitial value")
# Assign the Comp value of the next row to the current row in 'Comp_going'
results$Comp_going[zzz_rows] <- results$Comp[zzz_rows + 1]
results = results %>% mutate(Comp_from_to = paste(Comp,Comp_going,sep = "_"))
results2 = results %>% filter(!is.na(Comp_going))# Delete players that went somewhere else not top 5 legaues

#save as csv
#write.csv(results2, "csv data/results.csv", row.names = FALSE)
```

data for the positions before that
modify dataa
```{r}
dataa <- read_csv("csv data/results.csv")
dataa = dataa %>% filter(Squad == "zzzinitial value")

dataa = dataa %>% mutate(gb = ifelse(Transfer_Value <= 0, 0, 1)) %>% # If 0, then it is bad transfer
  filter(is.na(Transfer_Value) == FALSE) #put the NA values out
#-700000 is the median

cols_to_modify <- grep("pct|percent", names(dataa), ignore.case = TRUE)

# Replace NA values with 0 in identified columns (the one that are percentage)
dataa[cols_to_modify] <- lapply(dataa[cols_to_modify], function(x) {
  ifelse(is.na(x), 0, x)
})
```


```{r}
medians_table <- dataa %>%
  group_by(Age) %>%
  summarise(Median_Transfer_Value = median(Transfer_Value, na.rm = TRUE),Count = n())
# Take the median for each age of the Transfer_Value(my metric) and look if good or bad

# Add gb_age column (If a tranfer is succesfull or not)
dataa <- dataa %>%
  left_join(medians_table, by = "Age") %>%
  mutate(gb_age = ifelse(Transfer_Value > Median_Transfer_Value, 1, 0))

```

add pca to data
```{r}
dataa = left_join(dataa, pca_join_age_diff2total, by = c("player_url","Season_End_Year","gb_age"))
write.csv(dataa, "csv data/dataa.csv", row.names = FALSE)
```

data generation for single positions and all positions
```{r}
#data load it 
dataa <- read_csv("csv data/dataa123.csv")
dataa[1] <- NULL

mid_data <- dataa %>% 
  dplyr::filter(Pos %in% c("MF", "MF,FW", "DF,MF","MF,DF","FW,MF")) %>%#mid
  dplyr::select(-c(
    "global_transfer_index", "player_url", "Season_End_Year",
    "Squad", "Player", "Born","gb","Median_Transfer_Value","Is_Transfer","Transfer_Group","player_market_value_euro","MarketValue","rownr","squad","count","Count","Mn_per_Start_Starts",
    "Mn_per_Sub_Subs","Url","Pos"
  )) %>% dplyr::mutate(
    npxG_per_Sh_Expected = ifelse(is.na(npxG_per_Sh_Expected), 0, npxG_per_Sh_Expected),
    G_per_SoT_Standard = ifelse(is.na(G_per_SoT_Standard), 0, G_per_SoT_Standard),
    Dist_Standard = ifelse(is.na(Dist_Standard), 0, Dist_Standard),
    G_per_Sh_Standard = ifelse(is.na(G_per_Sh_Standard), 0, G_per_Sh_Standard)
  )

mid_data$Age <- as.numeric(mid_data$Age)
colnames(mid_data) <- make.names(colnames(mid_data), unique = TRUE)
mid_data <- na.omit(mid_data)#1 away after transformations

mid_data$gb_age <- as.factor(mid_data$gb_age)


df_data <- dataa %>% 
  dplyr::filter(Pos %in% c("DF", "DF,FW", "DF,FW","FW,DF","FW,DF")) %>%#df
  dplyr::select(-c(
    "global_transfer_index", "player_url", "Season_End_Year",
    "Squad", "Player", "Born","gb","Median_Transfer_Value","Is_Transfer","Transfer_Group","player_market_value_euro","MarketValue","rownr","squad","count","Count","Mn_per_Start_Starts",
    "Mn_per_Sub_Subs","Url","Pos"
  )) %>% dplyr::mutate(
    npxG_per_Sh_Expected = ifelse(is.na(npxG_per_Sh_Expected), 0, npxG_per_Sh_Expected),
    G_per_SoT_Standard = ifelse(is.na(G_per_SoT_Standard), 0, G_per_SoT_Standard),
    Dist_Standard = ifelse(is.na(Dist_Standard), 0, Dist_Standard),
    G_per_Sh_Standard = ifelse(is.na(G_per_Sh_Standard), 0, G_per_Sh_Standard)
  )
df_data$Age <- as.numeric(df_data$Age)
colnames(df_data) <- make.names(colnames(df_data), unique = TRUE)
df_data <- na.omit(df_data)
df_data$gb_age <- as.factor(df_data$gb_age)

fw_data <- dataa %>% 
  dplyr::filter(Pos %in% c("FW", "MF,FW", "DF,FW","FW,DF","FW,MF")) %>%#fw
  dplyr::select(-c(
    "global_transfer_index", "player_url", "Season_End_Year",
    "Squad", "Player", "Born","gb","Median_Transfer_Value","Is_Transfer","Transfer_Group","player_market_value_euro","MarketValue","rownr","squad","count","Count","Mn_per_Start_Starts",
    "Mn_per_Sub_Subs","Url","Pos"
  )) %>% dplyr::mutate(
    npxG_per_Sh_Expected = ifelse(is.na(npxG_per_Sh_Expected), 0, npxG_per_Sh_Expected),
    G_per_SoT_Standard = ifelse(is.na(G_per_SoT_Standard), 0, G_per_SoT_Standard),
    Dist_Standard = ifelse(is.na(Dist_Standard), 0, Dist_Standard),
    G_per_Sh_Standard = ifelse(is.na(G_per_Sh_Standard), 0, G_per_Sh_Standard)
  )
fw_data$Age <- as.numeric(fw_data$Age)
colnames(fw_data) <- make.names(colnames(fw_data), unique = TRUE)
fw_data <- na.omit(fw_data)
fw_data$gb_age <- as.factor(fw_data$gb_age)

#train test

train_data_fw <- fw_data %>%
  group_by(gb_age) %>%
  sample_frac(0.8)

test_data_fw  <- anti_join(fw_data, train_data_fw)

train_data_df <- df_data %>%
  group_by(gb_age) %>%
  sample_frac(0.8)

test_data_df<- anti_join(df_data, train_data_df)

train_data_mid <- mid_data %>%
  group_by(gb_age) %>%
  sample_frac(0.8)

 test_data_mid <- anti_join(mid_data, train_data_mid)

tot_data <- dataa %>% 
  dplyr::filter(Pos != "GK") %>%#no gk
  dplyr::select(-c(
    "global_transfer_index", "player_url", "Season_End_Year",
    "Squad", "Player", "Born","gb","Median_Transfer_Value","Is_Transfer","Transfer_Group","player_market_value_euro","MarketValue","rownr","squad","count","Count","Mn_per_Start_Starts",
    "Mn_per_Sub_Subs","Pos","Url"
  )) %>% dplyr::mutate(
    npxG_per_Sh_Expected = ifelse(is.na(npxG_per_Sh_Expected), 0, npxG_per_Sh_Expected),
    G_per_SoT_Standard = ifelse(is.na(G_per_SoT_Standard), 0, G_per_SoT_Standard),
    Dist_Standard = ifelse(is.na(Dist_Standard), 0, Dist_Standard),
    G_per_Sh_Standard = ifelse(is.na(G_per_Sh_Standard), 0, G_per_Sh_Standard)
  )

tot_data$Age <- as.numeric(tot_data$Age)
colnames(tot_data) <- make.names(colnames(tot_data), unique = TRUE)
tot_data <- na.omit(tot_data)#1 away after transformations

tot_data$gb_age <- as.factor(tot_data$gb_age)

train_data_tot <- tot_data %>%
  group_by(gb_age) %>%
  sample_frac(0.8)

test_data_tot <- anti_join(tot_data, train_data_tot)

```

rid of age for all datasets
```{r}
# Remove "Age" column from datasets
train_data_df2 <- train_data_df[, names(train_data_df) != "Age"]
test_data_df2 <- test_data_df[, names(test_data_df) != "Age"]

train_data_mid2 <- train_data_mid[, names(train_data_mid) != "Age"]
test_data_mid2 <- test_data_mid[, names(test_data_mid) != "Age"]

train_data_fw2 <- train_data_fw[, names(train_data_fw) != "Age"]
test_data_fw2 <- test_data_fw[, names(test_data_fw) != "Age"]

train_data_tot2 <- train_data_tot[, names(train_data_tot) != "Age"]
test_data_tot2 <- test_data_tot[, names(test_data_tot) != "Age"]
```

Save all important the datasets as CSV files
```{r}
# write.csv(train_data_df2, "csv data/train_data_df2.csv", row.names = FALSE)
# write.csv(test_data_df2, "csv data/test_data_df2.csv", row.names = FALSE)
# write.csv(test_data_df2, "csv data/test_data_fw2.csv", row.names = FALSE)
# write.csv(train_data_fw2, "csv data/train_data_fw2.csv", row.names = FALSE)
# write.csv(train_data_mid2, "csv data/train_data_mid2.csv", row.names = FALSE)
# write.csv(test_data_mid2, "csv data/test_data_mid2.csv", row.names = FALSE)
# write.csv(train_data_tot2, "csv data/train_data_tot2.csv", row.names = FALSE)
# write.csv(test_data_tot2, "csv data/test_data_tot2.csv", row.names = FALSE)
# #
# write.csv(df_data, "csv data/df_data.csv", row.names = FALSE)
# write.csv(mid_data, "csv data/mid_data.csv", row.names = FALSE)
# write.csv(fw_data, "csv data/fw_data.csv", row.names = FALSE)
# write.csv(tot_data, "csv data/tot_data.csv", row.names = FALSE)

```