---
title: "PCA"
author: "Justin Lampman"
date: "2025-01-22"
output: html_document
---

For the code to work without loading any data previously, you first have to run PCA.Rmd, then you have to run Daten.Rmd to have all data. To run the models you have to run rf.Rmd. For the plots you need to run plots.Rmd. 
To load all data without running the code, you can use the csv files in the csv data folder and run them in the second chunk. The models are saved in the models folder. The plots are saved in the plots folder. The data is saved in the data folder.

packages
```{r}
library(worldfootballR)
library(tidyverse)
library(ggrepel)
library(stringi)
library(stringr)
library(worldfootballR)
library(corrr)
library(ggcorrplot)
library(factoextra)
library(purrr)
library(vivid)
library(scales)
library(reshape2)  # For data transformation
library(viridis)
library(ranger)
library(Metrics)
mapping = read.csv("csv data/names.csv")# Mapping for clubs of transfermakt data and fbref data
```

If you do not want to run the code, you can also use the csv files in the csv data folder and load these to have all the data, to run the models and get the plots
```{r}
pca_join_age_diff2total <- read.csv("csv data/pca.csv")
dataa <- read_csv("csv data/dataa123.csv")%>% select(-1)
#
rfreg_tot_vivi2 <- read.csv("csv data/rfreg_tot_vivi2.csv")
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
data_butlook <- read_csv("csv data/data.csv")
#
mid_data2reg <- read.csv("csv data/mid_data2reg.csv")
df_data2reg <- read.csv("csv data/df_data2reg.csv")
tot_data2reg <- read.csv("csv data/vtot_data2reg.csv")
fw_data2reg <- read.csv("csv data/fw_data2reg.csv")
fw_data2bin <- read.csv("csv data/fw_data2bin.csv")
mid_data2bin <- read.csv("csv data/mid_data2bin.csv")
df_data2bin <- read.csv("csv data/df_data2bin.csv")
tot_data2bin <- read.csv("csv data/tot_data2bin.csv")
#
combined_varimp_bin2 <- read.csv("csv data/combined_varimp_bin2.csv")
combined_varimp_reg2 <- read.csv("csv data/combined_varimp_reg2.csv")
#
regbin <- read.csv("csv data/regbin.csv")
#
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
train_data_df2 <- read.csv("csv data/train_data_df2.csv")
test_data_df2 <- read.csv("csv data/test_data_df2.csv")
test_data_fw2 <- read.csv("csv data/test_data_fw2.csv") 
train_data_fw2 <- read.csv("csv data/train_data_fw2.csv")
train_data_mid2 <- read.csv("csv data/train_data_mid2.csv")
test_data_mid2 <- read.csv("csv data/test_data_mid2.csv")
train_data_tot2 <- read.csv("csv data/train_data_tot2.csv")
test_data_tot2 <- read.csv("csv data/test_data_tot2.csv")
#
df_data <- read.csv("csv data/df_data.csv")
mid_data <- read.csv("csv data/mid_data.csv")
fw_data <- read.csv("csv data/fw_data.csv")
tot_data <- read.csv("csv data/tot_data.csv")
#
rfbin_fw2 <- readRDS("models/rfbin_fw2.rds")
rfbin_df2 <- readRDS("models/rfbin_df2.rds")
rfbin_mid2 <- readRDS("models/rfbin_mid2.rds")
rfbin_tot2 <- readRDS("models/rfbin_tot2.rds")
rfreg_fw2 <- readRDS("models/rfreg_fw2.rds")
rfreg_df2 <- readRDS("models/rfreg_df2.rds")
rfreg_mid2 <- readRDS("models/rfreg_mid2.rds")
rfreg_tot2 <- readRDS("models/rfreg_tot2.rds")
# load these for the data from Transfermarkt and Fbref
all_val <- read.csv("csv data/all_val.csv")
final_combined_data <- read.csv("csv data/final_combined_data.csv")
joined_data = read.csv("csv data/team_data_pca.csv")
```

data for teams
```{r}
big5_team_standard <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("standard"), team_or_player= "team")

big5_team_shooting <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("shooting"), team_or_player= "team")

big5_team_passing <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("passing"), team_or_player= "team")

big5_team_passing_types <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("passing_types"), team_or_player= "team")

big5_team_gca <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("gca"), team_or_player= "team")

big5_team_defense <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("defense"), team_or_player= "team")

big5_team_possession <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("possession"), team_or_player= "team")

big5_team_playing_time <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("playing_time"), team_or_player= "team")

big5_team_misc <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("misc"), team_or_player= "team")

big5_team_keepers <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("keepers"), team_or_player= "team")

big5_team_keepers_adv <- fb_big5_advanced_season_stats(season_end_year= c(2018:2024), stat_type= c("keepers_adv"), team_or_player= "team")

```

team data join and format
```{r}
datasets <- list(
  big5_team_standard,
  big5_team_shooting,
  big5_team_passing,
  big5_team_passing_types,
  big5_team_gca,
  big5_team_defense,
  big5_team_possession,
  big5_team_playing_time,
  big5_team_misc,
  big5_team_keepers,
  big5_team_keepers_adv
)

# Join all datasets by the first 4 columns
joined_data <- reduce(datasets, function(x, y) {
  full_join(x, y, by = colnames(x)[1:4])
})

joined_data <- joined_data %>% 
  dplyr::select(!matches("\\.x$|\\.y$"))
#save
#write.csv(joined_data, "csv data/team_data_pca.csv")
#joined_data = read.csv("csv data/team_data_pca.csv")
```

team data join and format 2
```{r}
teams = joined_data %>% dplyr::select("Season_End_Year","Squad","Team_or_Opponent","PrgC_Progression","PrgP_Progression",
       "Sh_Standard","Dist_Standard",
       "Att_Total","TotDist_Total","PrgDist_Total","Att_Medium","Att_Short","Att_Long","KP","Final_Third","PPA","CrsPA","PrgP",
       "TB_Pass","Sw_Pass","Crs_Pass",
       "PassLive_SCA","PassDead_SCA","Def_SCA","TO_GCA",
       "Def.3rd_Tackles","Mid.3rd_Tackles","Att.3rd_Tackles","Att_Challenges","Blocks_Blocks","Tkl.Int","Clr",
       "Def.Pen_Touches","Def.3rd_Touches","Mid.3rd_Touches","Att.3rd_Touches","Att.Pen_Touches","TotDist_Carries",                 "PrgDist_Carries","PrgC_Carries","Final_Third_Carries","CPA_Carries","PrgR_Receiving","Att_Take",
       "Won_Aerial","Lost_Aerial","Fls","Fld","Off") %>%
  dplyr::filter(Team_or_Opponent == "team")

teams$Squad <- str_replace_all(teams$Squad, " ", "_")

teams = teams %>%
  mutate(Squad_Season = str_c(str_replace_all(Squad, " ", "_"), "_", Season_End_Year)) 

rownames(teams) <- teams$Squad_Season

teams = teams %>% dplyr::select(-c("Season_End_Year","Squad","Team_or_Opponent","Squad_Season"))
```

values for the pcas
```{r}
euro_scaled <- scale(teams)
projection_matrix = eigen(cor(teams))$vectors[, 1:4]
projected_x <- euro_scaled %*% projection_matrix
colnames(projected_x) <- c("PC1", "PC2","PC3","PC4")
head(projected_x)
```

plots for pcas by team
dont need but keep it
```{r}
# Convert to data frame and extract team names and years
df_pca <- as.data.frame(projected_x)
df_pca$Team <- gsub("_\\d{4}", "", rownames(projected_x))
df_pca$Year <- as.numeric(gsub(".*_(\\d{4})", "\\1", rownames(projected_x)))

# Reshape data to long format
df_long <- df_pca %>%
  pivot_longer(cols = starts_with("PC"), names_to = "Component", values_to = "Value")

# Choose specific teams to visualize
selected_teams <- c("Atlético_Madrid", "Barcelona", "Liverpool", "Leicester_City", "Manchester_City")

# Filter the data for teams
df_filtered <- df_long %>%
  filter(Team %in% selected_teams)

# Plot the PCA values over time for some teams. Not in Thesis
ggplot(df_filtered, aes(x = Year, y = Value, color = Team, group = Team)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Component, scales = "free_y") +
  labs(title = "PCA Values Over Time for Selected Teams",
       x = "Year",
       y = "PCA Value") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

make pca bind to each player with all pcas 
```{r}
data_butlook <- read_csv("data.csv")#read csv from Daten.Rmd to combine playe with pca from teams
pca_team_before <- dataa %>%    
  mutate(Squad_Season = str_c(str_replace_all(squad, " ", "_"), "_", Season_End_Year)) %>% 
  left_join((as.data.frame(projected_x) %>%
  rownames_to_column(var = "Team")), by = c("Squad_Season" = "Team")) %>%    
  dplyr::select(squad, player_url, Squad_Season, gb_age, Season_End_Year,PC1,PC2,PC3,PC4)


# Step 2: Get filtered URLs
filtered_urls_pca_diff <- unique(pca_team_before$player_url)

inbetween_pca_diff = pca_team_before %>% dplyr::select(player_url,Season_End_Year) %>%
  mutate(Season_End_Year = Season_End_Year + 1)

#adjust data
data_butlook = data_butlook %>%
  arrange(player_url, Season_End_Year, rownr)

# Perform the join as before
pca_team_after <- inner_join(inbetween_pca_diff, data_butlook, 
                   by = c("player_url" = "player_url", "Season_End_Year" = "Season_End_Year")) %>%
  dplyr::select(player_url, Season_End_Year, Squad, rownr) %>%
  group_by(player_url, Season_End_Year) %>%
  slice_min(rownr) %>%
  ungroup() %>%
  mutate(Squad_Season = str_c(str_replace_all(Squad, " ", "_"), "_", Season_End_Year ),
         Season_End_Year = Season_End_Year)%>%
left_join((as.data.frame(projected_x) %>%
  rownames_to_column(var = "Team")), by = c("Squad_Season" = "Team")) %>%    
  dplyr::select(Squad, player_url, Squad_Season, Season_End_Year,PC1,PC2,PC3,PC4) %>% 
  mutate(Season_End_Year = Season_End_Year - 1) %>%
  mutate(PC1a = PC1,PC2a = PC2,PC3a = PC3,PC4a = PC4) %>%
  dplyr::select(-PC1,-PC2,-PC3,-PC4)

```

total differences between pcas
```{r}
pca_join_age_diff2total <- inner_join(pca_team_before, pca_team_after, by = c("player_url", "Season_End_Year")) %>% 
  # Calculate the differences
  mutate(
    PC1_diff = PC1 - PC1a,
    PC2_diff = PC2 - PC2a,
    PC3_diff = PC3 - PC3a,
    PC4_diff = PC4 - PC4a,
    total_diff = abs(PC1_diff) + abs(PC2_diff) + abs(PC3_diff) + abs(PC4_diff)
  ) %>% 
  # Drop unnecessary columns
  dplyr::select(
    -PC1, -PC2, -PC3, -PC4,
    -PC1a, -PC2a, -PC3a, -PC4a,
    -Squad_Season.x, -Squad_Season.y,
    -Squad, -squad
  )
```

save the pca file
```{r}
# write.csv(pca_join_age_diff2total, "csv data/pca.csv")
```

