---
title: "Plots"
author: "Justin Lampman"
date: "2025-01-22"
output: html_document
---

For the code to work without loading any data previously, you first have to run PCA.Rmd, then you have to run Daten.Rmd to have all data. To run the models you have to run rf.Rmd. For the plots you need to run plots.Rmd. 
To load all data without running the code, you can use the csv files in the csv data folder and run them in the second chunk. The models are saved in the models folder. The plots are saved in the plots folder. The data is saved in the data folder.

pakages
```{r}
library(worldfootballR)
library(tidyverse)
library(ggrepel)
library(stringi)
library(stringr)
library(worldfootballR)
library(corrr)
library(ggcorrplot)
library(factoextra)
library(purrr)
library(vivid)
library(reshape2)  # For data transformation
library(viridis)
library(scales)#for the Age_vs_Value.png plot (not scientific)
library(ranger)
library(Metrics)
mapping = read.csv("csv data/names.csv")# Mapping for clubs of transfermakt data and fbref data
```

If you do not want to run the code, you can also use the csv files in the csv data folder and load these to have all the data, to run the models and get the plots
```{r}
pca_join_age_diff2total <- read.csv("csv data/pca.csv")
dataa <- read_csv("csv data/dataa123.csv")%>% select(-1)
#
rfreg_tot_vivi2 <- read.csv("csv data/rfreg_tot_vivi2.csv")
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
data_butlook <- read_csv("csv data/data.csv")
#
mid_data2reg <- read.csv("csv data/mid_data2reg.csv")
df_data2reg <- read.csv("csv data/df_data2reg.csv")
tot_data2reg <- read.csv("csv data/vtot_data2reg.csv")
fw_data2reg <- read.csv("csv data/fw_data2reg.csv")
fw_data2bin <- read.csv("csv data/fw_data2bin.csv")
mid_data2bin <- read.csv("csv data/mid_data2bin.csv")
df_data2bin <- read.csv("csv data/df_data2bin.csv")
tot_data2bin <- read.csv("csv data/tot_data2bin.csv")
#
combined_varimp_bin2 <- read.csv("csv data/combined_varimp_bin2.csv")
combined_varimp_reg2 <- read.csv("csv data/combined_varimp_reg2.csv")
#
regbin <- read.csv("csv data/regbin.csv")
#
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
train_data_df2 <- read.csv("csv data/train_data_df2.csv")
test_data_df2 <- read.csv("csv data/test_data_df2.csv")
test_data_fw2 <- read.csv("csv data/test_data_fw2.csv") 
train_data_fw2 <- read.csv("csv data/train_data_fw2.csv")
train_data_mid2 <- read.csv("csv data/train_data_mid2.csv")
test_data_mid2 <- read.csv("csv data/test_data_mid2.csv")
train_data_tot2 <- read.csv("csv data/train_data_tot2.csv")
test_data_tot2 <- read.csv("csv data/test_data_tot2.csv")
#
df_data <- read.csv("csv data/df_data.csv")
mid_data <- read.csv("csv data/mid_data.csv")
fw_data <- read.csv("csv data/fw_data.csv")
tot_data <- read.csv("csv data/tot_data.csv")
#
rfbin_fw2 <- readRDS("models/rfbin_fw2.rds")
rfbin_df2 <- readRDS("models/rfbin_df2.rds")
rfbin_mid2 <- readRDS("models/rfbin_mid2.rds")
rfbin_tot2 <- readRDS("models/rfbin_tot2.rds")
rfreg_fw2 <- readRDS("models/rfreg_fw2.rds")
rfreg_df2 <- readRDS("models/rfreg_df2.rds")
rfreg_mid2 <- readRDS("models/rfreg_mid2.rds")
rfreg_tot2 <- readRDS("models/rfreg_tot2.rds")
# load these for the data from Transfermarkt and Fbref
all_val <- read.csv("csv data/all_val.csv")
final_combined_data <- read.csv("csv data/final_combined_data.csv")
joined_data = read.csv("csv data/team_data_pca.csv")
```

generate_pdp helper function
```{r}
generate_pdp <- function(model, data, features, grid_resolution = 20) {
  
  # Custom predict function (for models that require type = "prob")
  predict_rf <- function(object, newdata, ...) {
    newdata <- as.data.frame(newdata)
    predict(object, newdata = newdata, type = "prob")[, 2]
  }

  # Compute PDPs for each feature
  pdp_data <- lapply(features, function(var) {
    partial(
      object = model,
      pred.var = var,
      train = data,
      grid.resolution = grid_resolution,
      .f = predict_rf,
      progress = "text"
    ) %>%
      as.data.frame() %>%
      rename(value = 1, yhat = 2) %>%
      mutate(value = as.numeric(value), Feature = var)
  })

  # Combine all PDP results into a single dataframe
  pdp_df <- bind_rows(pdp_data)

  # Plot 
  ggplot(pdp_df, aes(x = value, y = yhat)) +
    geom_line(linewidth = 1, color = "blue") +
    labs(
      title = "",
      x = "Feature Value",
      y = "Predicted Response"
    ) +
    facet_wrap(~ Feature, scales = "free_x") +
    theme_minimal(base_size = 22) +
    theme(
      axis.text.x = element_text(size = 20),
      axis.text.y = element_text(size = 20),
      strip.text = element_text(size = 22),
      plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
      axis.title.x = element_text(size = 22),
      axis.title.y = element_text(size = 22)
    )
}
```

generate_pdp_two_vars helper function
```{r}
generate_pdp_two_vars <- function(model, data, feature1, feature2, grid_resolution = 20) {
  
  # Custom predict function to get class 1 probability
  predict_rf <- function(object, newdata, ...) {
    newdata <- as.data.frame(newdata)
    predict(object, newdata = newdata, type = "prob")[, 2]
  }

  # Compute 2D partial dependence data
  pdp_data <- partial(
    object = model,
    pred.var = c(feature1, feature2),
    train = data,
    grid.resolution = grid_resolution,
    .f = predict_rf,
    progress = "text"
  ) %>%
    as.data.frame() %>%
    rename(value1 = 1, value2 = 2, yhat = 3)

  # Plot the 2D PDP
  ggplot(pdp_data, aes(x = value1, y = value2, fill = yhat)) +
    geom_tile() +
    scale_fill_gradient2(low = "red", high = "blue", mid = "lightblue", midpoint = 0) +
    labs(
      title = "",
      x = feature1,
      y = feature2,
      fill = "Predicted Response"
    ) +
    theme_minimal(base_size = 22) +
    theme(
      axis.text.x = element_text(size = 24),
      axis.text.y = element_text(size = 24),
      plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
      axis.title.x = element_text(size = 26),
      axis.title.y = element_text(size = 26),
      legend.text = element_text(size = 20),
      legend.title = element_text(size = 26)
    )
}
```

generate_pdp_not_nice helper function
```{r}
generate_pdp_not_nice <- function(model, data, features, grid_resolution = 20) {
  
  # Custom predict function (for models that require type = "prob")
  predict_rf <- function(object, newdata, ...) {
    newdata <- as.data.frame(newdata)
    predict(object, newdata = newdata, type = "prob")[, 2]
  }

  # Compute PDPs for each feature
  pdp_data <- lapply(features, function(var) {
    partial(
      object = model,
      pred.var = var,
      train = data,
      grid.resolution = grid_resolution,
      .f = predict_rf,
      progress = "text"
    ) %>%
      as.data.frame() %>%
      rename(value = 1, yhat = 2) %>%
      mutate(value = as.numeric(value), Feature = var)
  })

  # Combine all PDP results into a single dataframe
  pdp_df <- bind_rows(pdp_data)

  # Plot PDP using ggplot2 with larger text
  ggplot(pdp_df, aes(x = value, y = yhat)) +  
    geom_line(linewidth = 1, color = "blue") +  
    #geom_rug(aes(x = value), sides = "b", alpha = 0.3) +  
    #geom_rug(aes(y = yhat), sides = "l", alpha = 0.3) +  
    labs(
      title = "",  
      x = "Feature Value",  
      y = "Predicted Response"  
    ) +  
    facet_wrap(~ Feature, scales = "free_x") +  
    theme_minimal()
}
```

plot for age and value
```{r}
#from RF 
ggplot(dataa, aes(x = factor(Age), y = Transfer_Value)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "",
       x = "Age",
       y = "Transfer Value") +
  theme_minimal() +
  scale_y_continuous(
    limits = c(-1e7, 1e7),
    labels = comma                # formats with commas, no sci notation
  )

ggsave("Plots/Figure1.png")
```

bar plot pca
```{r}
# Create the data frame manually
df_pca = as.data.frame(projected_x[c("Eibar_2019", "Manchester_City_2021", "Sevilla_2022", "Atlético_Madrid_2018","Dortmund_2023"), ])

df_pca$Team <- rownames(df_pca)

# Convert data to long format for ggplot
df_long <- df_pca %>%
  pivot_longer(cols = starts_with("PC"), names_to = "Component", values_to = "Value")


# Adjust theme for larger text and legend position
plot_pca = ggplot(df_long, aes(x = gsub("_", " ", Team), y = Value, fill = Component)) +   
  geom_bar(stat = "identity", position = "dodge") +   
  labs(title = "Comparison of PCA Components for Teams",        
       x = "Team",        
       y = "PCA Value") +   
  theme_minimal(base_size = 22) +  # Increase overall text size
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 20),  # Bigger x-axis labels
    axis.text.y = element_text(size = 20),  # Bigger y-axis labels
    legend.text = element_text(size = 20),  # Bigger legend text
    legend.title = element_text(size = 22),  # Bigger legend title
    plot.title = element_text(size = 24, hjust = 0.5),  # Centered & bigger title
    legend.position = "bottom",  # Move legend below the plot
    legend.key.size = unit(1.5, "cm")  # Make legend keys bigger
  )

# Export with higher resolution, adjusted size, and legend-friendly layout
ggsave(filename = "Plots/Figure4.png", plot = plot_pca, 
       width = 8, height = 10, dpi = 300, scale = 1.5)

```

plot for pdp whole model bin
```{r}
a99 = generate_pdp(rfbin_tot2, tot_data[ , names(tot_data) != "gb_age"] , c("PC2_diff","PC1_diff","PC3_diff","PC4_diff","total_diff","PPM_Team.Success","xGplus_per__minus__Team.Success..xG","xGplus_per__minus_90_Team.Success..xG","On_minus_Off_Team.Success"))

# Define readable labels for the features
feature_labels <- c(
  "PC2_diff" = "PC2 Difference",
  "PC1_diff" = "PC1 Difference",
  "PC3_diff" = "PC3 Difference",
  "PC4_diff" = "PC4 Difference",
  "total_diff" = "Total PC Difference",
  "PPM_Team.Success" = "Points per Match",
  "xGplus_per__minus__Team.Success..xG" = "Net Expected Goals",
  "xGplus_per__minus_90_Team.Success..xG" = "xG for - against per90",
  "On_minus_Off_Team.Success" = "Impact On vs Off Pitch"
)

# Modify facet labels using labeller
a99 = a99 + facet_wrap(~ Feature, scales = "free_x", labeller = as_labeller(feature_labels))


ggsave("Plots/Figure5.png", a99, width = 8, height = 10, dpi = 300, scale = 1.5)
```

interactions for whole model bin
```{r}
# Extract variable importance from the trained model
var_importance <- rfbin_tot2$variable.importance

# Convert to a dataframe for sorting
var_importance_df <- data.frame(Variable = names(var_importance), Importance = var_importance)

# Sort by importance in descending order
var_importance_df <- var_importance_df[order(-var_importance_df$Importance), ]

# Get the top 40 most important variables
top_variables <- var_importance_df$Variable[1:40]

# delete the first col and make it the names
rfbin_tot_vivi2 <- as.data.frame(rfbin_tot_vivi2)
rownames(rfbin_tot_vivi2) <- rfbin_tot_vivi2[[1]]
rfbin_tot_vivi2 <- rfbin_tot_vivi2[, -1]

# Subset and reorder rfbin_tot_vivi2 based on the sorted variable importance
sorted_mat <- rfbin_tot_vivi2[top_variables, top_variables]
sorted_mat <- as.matrix(sorted_mat)

# Plot the heatmap
vivi1 = viviHeatmap(mat = sorted_mat,angle = 45)

ggsave("Plots/Figure6.png", vivi1, width = 10, height = 10, units = "in", dpi = 300)
```

pdp for fw model
```{r}
pdp_fw = generate_pdp(rfbin_fw2, fw_data2bin , c("PC4_diff","PC1_diff","PC2_diff","Succ_percent_Take","total_diff", "PC3_diff"))#top 6 variables for fw

# Define readable labels for the features
feature_labels2 <- c(
  "PC2_diff" = "PC2 Difference",
  "PC1_diff" = "PC1 Difference",
  "PC3_diff" = "PC3 Difference",
  "PC4_diff" = "PC4 Difference",
  "total_diff" = "Total PC Difference",
  "Succ_percent_Take" = "Successful Take-On %"
)

# Modify facet labels using labeller
pdp_fw = pdp_fw + facet_wrap(~ Feature, scales = "free_x", labeller = as_labeller(feature_labels2))

ggsave("Plots/Figure9.png", pdp_fw, width = 8, height = 10, dpi = 300, scale = 1.5)

```

interactions for fw model
```{r}
 # Extract variable importance from the trained model
var_importancefw <- rfbin_fw2$variable.importance

# Convert to a dataframe for sorting
var_importance_dffw <- data.frame(Variable = names(var_importancefw), Importance = var_importancefw)

# Sort by importance in descending order
var_importance_dffw <- var_importance_dffw[order(-var_importance_dffw$Importance), ]

# Get the top 40 most important variables
top_variablesfw <- var_importance_dffw$Variable[1:40]

# delete the first col and make it the names
rfbin_fw_vivi2 <- as.data.frame(rfbin_fw_vivi2)
rownames(rfbin_fw_vivi2) <- rfbin_fw_vivi2[[1]]
rfbin_fw_vivi2 <- rfbin_fw_vivi2[, -1]

# Subset and reorder rfbin_fw_vivi2 based on the sorted variable importance
sorted_matfw <- rfbin_fw_vivi2[top_variablesfw, top_variablesfw]
sorted_matfw <- as.matrix(sorted_matfw)

# Plot the heatmap
vivi2 = viviHeatmap(mat = sorted_matfw,angle = 45)

ggsave("Plots/Figure10.png", vivi2, width = 10, height = 10, units = "in", dpi = 300)
```

interactions for mid model
```{r}
 # Extract variable importance from the trained model
var_importancemid <- rfbin_mid2$variable.importance

# Convert to a dataframe for sorting
var_importance_dfmid <- data.frame(Variable = names(var_importancemid), Importance = var_importancemid)

# Sort by importance in descending order
var_importance_dfmid <- var_importance_dfmid[order(-var_importance_dfmid$Importance), ]

# Get the top 40 most important variables
top_variablesmid <- var_importance_dfmid$Variable[1:40]

# delete the first col and make it the names
rfbin_mid_vivi2 <- as.data.frame(rfbin_mid_vivi2)
rownames(rfbin_mid_vivi2) <- rfbin_mid_vivi2[[1]]
rfbin_mid_vivi2 <- rfbin_mid_vivi2[, -1]

# Subset and reorder rfbin_tot_vivi2 based on the sorted variable importance
sorted_matmid <- rfbin_mid_vivi2[top_variablesmid, top_variablesmid]
sorted_matmid <- as.matrix(sorted_matmid)

# Plot the heatmap
vivi3 = viviHeatmap(mat = sorted_matmid,angle = 45)

ggsave("Plots/Figure14.png", vivi3, width = 10, height = 10, units = "in", dpi = 300)
```
#### muss noch machen

interactions for df model
```{r}
  # Extract variable importance from the trained model
var_importancedf <- rfbin_df2$variable.importance

# Convert to a dataframe for sorting
var_importance_dfdf <- data.frame(Variable = names(var_importancedf), Importance = var_importancedf)

# Sort by importance in descending order
var_importance_dfdf <- var_importance_dfdf[order(-var_importance_dfdf$Importance), ]

# Get the top 40 most important variables
top_variablesdf <- var_importance_dfdf$Variable[1:40]

# delete the first col and make it the names
rfbin_df_vivi2 <- as.data.frame(rfbin_df_vivi2)
rownames(rfbin_df_vivi2) <- rfbin_df_vivi2[[1]]
rfbin_df_vivi2 <- rfbin_df_vivi2[, -1]

# Subset and reorder rfbin_tot_vivi2 based on the sorted variable importance
sorted_matdf <- rfbin_df_vivi2[top_variablesdf, top_variablesdf]
sorted_matdf <- as.matrix(sorted_matdf)

# Plot the heatmap
vivi4 = viviHeatmap(mat = sorted_matdf,angle = 45)

ggsave("Plots/Figure18.png", vivi4, width = 10, height = 10, units = "in", dpi = 300)
```

pdp for df model
```{r}
pdp_df = generate_pdp(rfbin_df2, df_data2bin , c("PC4_diff","PC1_diff","PC2_diff","Won_percent_Aerial","total_diff", "PC3_diff"))#top 6 variables for fw

# Define readable labels for the features
feature_labels3 <- c(
  "PC2_diff" = "PC2 Difference",
  "PC1_diff" = "PC1 Difference",
  "PC3_diff" = "PC3 Difference",
  "PC4_diff" = "PC4 Difference",
  "total_diff" = "Total PC Difference",
  "Won_percent_Aerial" = "Aerials Won %"
)

# Modify facet labels using labeller
pdp_df = pdp_df + facet_wrap(~ Feature, scales = "free_x", labeller = as_labeller(feature_labels3))

ggsave("Plots/Figure17.png", pdp_df, width = 8, height = 10, dpi = 300, scale = 1.5)
```

mid pdp
```{r}
pdp_mid = generate_pdp(rfbin_mid2, mid_data2bin , c("PC4_diff","PC1_diff","PC2_diff","xGplus_per__minus_90_Team.Success..xG","total_diff", "PC3_diff"))#top 6 variables for mid


# Define readable labels for the features
feature_labels4 <- c(
  "PC2_diff" = "PC2 Difference",
  "PC1_diff" = "PC1 Difference",
  "PC3_diff" = "PC3 Difference",
  "PC4_diff" = "PC4 Difference",
  "total_diff" = "Total PC Difference",
  "xGplus_per__minus_90_Team.Success..xG" = "xG for - against per90"
)

# Modify facet labels using labeller
pdp_mid = pdp_mid + facet_wrap(~ Feature, scales = "free_x", labeller = as_labeller(feature_labels4))

ggsave("Plots/Figure13.png", pdp_mid, width = 8, height = 10, dpi = 300, scale = 1.5)
```

tot interactions
```{r}
tot_int = generate_pdp_two_vars(rfbin_tot2, tot_data2bin, "xGplus_per__minus__Team.Success..xG", "PC2_diff")+
  labs(x = "xG for - against",y = "PC2 Difference")

ggsave("Plots/Figure7.png", tot_int, width = 10, height = 10, units = "in", dpi = 300)

```

fw interactions
```{r}
fw_int1 = generate_pdp_two_vars(rfbin_fw2, fw_data2bin, "Succ_percent_Take", "total_diff")+labs(x = "Successful Take-On %",y = "Total PC Difference")

ggsave("Plots/Figure11.png", fw_int1, width = 10, height = 10, units = "in", dpi = 300)

fw_int2 = generate_pdp_two_vars(rfbin_fw2, fw_data2bin, "Succ_percent_Take","Tkld_percent_Take")+labs(y = "Tackles Won %",x = "Successful Take-On %")

ggsave("Plots/Figure12.png", fw_int2, width = 10, height = 10, units = "in", dpi = 300)

```

mid interactions
```{r}
mid_int1 = generate_pdp_two_vars(rfbin_mid2, mid_data2bin, "Dead_Pass", "Ast_Per")+labs(x = "Dead Passes",y = "Assists per 90")

ggsave("Plots/Figure15.png", mid_int1, width = 10, height = 10, units = "in", dpi = 300)

mid_int2 = generate_pdp_two_vars(rfbin_mid2, mid_data2bin, "Dead_Pass", "Won_percent_Aerial")+labs(y = "Aerials Won %",x = "Dead Passes")

ggsave("Plots/Figure16.png", mid_int2, width = 10, height = 10, units = "in", dpi = 300)

```

df interctions
```{r}
df_int1 = generate_pdp_two_vars(rfbin_df2, df_data2bin,"Won_percent_Aerial",  "xGplus_per__minus_90_Team.Success..xG") + labs(y = "xG for - against per90",x = "Aerials Won %")

ggsave("Plots/Figure19.png", df_int1, width = 10, height = 10, units = "in", dpi = 300)
#then the same but for xGplus_per__minus_Team.Success..xG 

#then xGplus_per__minus__Team.Success..xG and xGplus_per__minus_90_Team.Success..xG which is discussed in one

```

comp from to and expected goals
```{r}
data_comptofrom = tot_data2bin
data_comptofrom$Comp_from_to <- as.factor(data_comptofrom$Comp_from_to)

one1 <- partial(object = rfbin_tot2,
                       pred.var = c("Comp_from_to", "xGplus_per__minus__Team.Success..xG"),
                       train = data_comptofrom,
                       grid.resolution = 10) %>%
    as.data.frame() %>%
    rename(value1 = 1, value2 = 2, yhat = 3)

generate_pdp_selected_lines_plot <- function(data) {
  
  # Make sure data in the correct format
  data <- data %>%
    mutate(value1 = as.factor(value1))  # Convert value to a factor for filtering
  
  # Filter for specific value categories
  data_filtered <- data %>%
    filter(value1 %in% c("Premier League_Serie A", "Bundesliga_Bundesliga"))%>% #Highest and lowest vals
    mutate(value1 = gsub("_", " to ", value1))
  
  # Plot with only selected categories colored
  ggplot(data, aes(x = value2, y = yhat, group = value1)) +
    geom_line(data = data, aes(x = value2, y = yhat), color = "gray", alpha = 0.4) +
    geom_line(data = data_filtered, aes(color = value1), linewidth = 1) +
    labs(
      title = "",
      x = "Net Expected Goals",
      y = "Predicted Response",
      color = "Competition From and To"
    ) +
    theme_minimal(base_size = 22) + 
    theme(
      axis.text.x = element_text(size = 24),  
      axis.text.y = element_text(size = 24),  
      plot.title = element_text(size = 26, hjust = 0.5, face = "bold"),  
      axis.title.x = element_text(size = 26),  
      axis.title.y = element_text(size = 26),  
      legend.text = element_text(size = 24),  
      legend.title = element_text(size = 26)
    )  
}

comp_xg = generate_pdp_selected_lines_plot(one1)

ggsave("Plots/Figure8.png", comp_xg, width = 10, height = 10, dpi = 300, scale = 1.5)
```

```{r}
top_20_features_tot_bin2 <- combined_varimp_bin2 %>%
  arrange(Rank_Avg) %>%
  head(22) %>%
  filter(!Variable %in% c("Comp_from_to", "Nation")) %>% #nation, comptofrom category out because not numeric
  pull(Variable)

top_20_features_fw_bin2 <- combined_varimp_bin2 %>%
  arrange(Rank_fw) %>%
  head(21) %>%
  filter(!Variable %in% c("Nation")) %>% #nation category out because not numeric
  pull(Variable)

top_20_features_mid_bin2 <- combined_varimp_bin2 %>%
  arrange(Rank_mid) %>%
  head(22) %>%
  filter(!Variable %in% c("Comp_from_to", "Nation")) %>% #nation, comptofrom category out because not numeric
  pull(Variable)

top_20_features_df_bin2 <- combined_varimp_bin2 %>%
  arrange(Rank_df) %>%
  head(22) %>%
  filter(!Variable %in% c("Comp_from_to", "Nation")) %>% #nation, comptofrom category out because not numeric
  pull(Variable)

top_20_features_tot_reg2 <- combined_varimp_reg2 %>%
  arrange(Rank_Avg) %>%
  head(21) %>%
  filter(!Variable %in% c("Nation")) %>% #nation category out because not numeric
  pull(Variable)

top_20_features_fw_reg2 <- combined_varimp_reg2 %>%
  arrange(Rank_fw) %>%
  head(20) %>%
  pull(Variable)

top_20_features_mid_reg2 <- combined_varimp_reg2 %>%
  arrange(Rank_mid) %>%
  head(21) %>%
  filter(!Variable %in% c("Nation")) %>% #nation category out because not numeric
  pull(Variable)

top_20_features_df_reg2 <- combined_varimp_reg2 %>%
  arrange(Rank_df) %>%
  head(20) %>%
  pull(Variable)

# Call the function
a12 = generate_pdp_not_nice(rfbin_tot2, tot_data[ , names(tot_data) != "gb_age"] , top_20_features_tot_bin2)
a22 = generate_pdp_not_nice(rfbin_fw2, fw_data[ , names(fw_data) != "gb_age"] , top_20_features_fw_bin2)
a32 = generate_pdp_not_nice(rfbin_mid2, mid_data[ , names(mid_data) != "gb_age"] , top_20_features_mid_bin2)
a42 = generate_pdp_not_nice(rfbin_df2, df_data[ , names(df_data) != "gb_age"] , top_20_features_df_bin2)
a52 = generate_pdp_not_nice(rfreg_tot2, tot_data[-1] , top_20_features_tot_reg2) + labs(y = "Predicted Transfer Value in Euro")
a62 = generate_pdp_not_nice(rfreg_fw2, fw_data[-1] , top_20_features_fw_reg2)+ labs(y = "Predicted Transfer Value in Euro")
a72 = generate_pdp_not_nice(rfreg_mid2, mid_data[-1] , top_20_features_mid_reg2)+ labs(y = "Predicted Transfer Value in Euro")
a82 = generate_pdp_not_nice(rfreg_df2, df_data[-1] , top_20_features_df_reg2)+ labs(y = "Predicted Transfer Value in Euro")

ggsave("Plots/Figure20.png", a12, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure21.png", a22, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure22.png", a32, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure23.png", a42, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure24.png", a52, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure25.png", a62, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure26.png", a72, width = 10, height = 10, units = "in", dpi = 300)
ggsave("Plots/Figure27.png", a82, width = 10, height = 10, units = "in", dpi = 300)
```