---
title: "Random Forest"
author: "Justin Lampman"
date: "2025-02-19"
output: html_document
---

For the code to work without loading any data previously, you first have to run PCA.Rmd, then you have to run Daten.Rmd to have all data. To run the models you have to run rf.Rmd. For the plots you need to run plots.Rmd.

To load all data without running the code, you can use the csv files in the csv data folder and run them in the second chunk. The models are saved in the models folder. The plots are saved in the plots folder. The data is saved in the data folder.

Packages
```{r}
library(worldfootballR)
library(tidyverse)
library(ggrepel)
library(stringi)
library(stringr)
library(worldfootballR)
library(corrr)
library(ggcorrplot)
library(factoextra)
library(purrr)
library(vivid)
library(scales)
library(reshape2)  # For data transformation
library(viridis)
library(ranger)
library(Metrics)
mapping = read.csv("csv data/names.csv")# Mapping for clubs of transfermakt data and fbref data
```

If you do not want to run the code, you can also use the csv files in the csv data folder and load these to have all the data, to run the models and get the plots
```{r}
pca_join_age_diff2total <- read.csv("csv data/pca.csv")
dataa <- read_csv("csv data/dataa123.csv")%>% select(-1)
#
rfreg_tot_vivi2 <- read.csv("csv data/rfreg_tot_vivi2.csv")
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
data_butlook <- read_csv("csv data/data.csv")
#
mid_data2reg <- read.csv("csv data/mid_data2reg.csv")
df_data2reg <- read.csv("csv data/df_data2reg.csv")
tot_data2reg <- read.csv("csv data/vtot_data2reg.csv")
fw_data2reg <- read.csv("csv data/fw_data2reg.csv")
fw_data2bin <- read.csv("csv data/fw_data2bin.csv")
mid_data2bin <- read.csv("csv data/mid_data2bin.csv")
df_data2bin <- read.csv("csv data/df_data2bin.csv")
tot_data2bin <- read.csv("csv data/tot_data2bin.csv")
#
combined_varimp_bin2 <- read.csv("csv data/combined_varimp_bin2.csv")
combined_varimp_reg2 <- read.csv("csv data/combined_varimp_reg2.csv")
#
regbin <- read.csv("csv data/regbin.csv")
#
rfbin_tot_vivi2 <- read.csv("csv data/rfbin_tot_vivi2.csv")
rfreg_fw_vivi2 <- read.csv("csv data/rfreg_fw_vivi2.csv")
rfbin_fw_vivi2 <- read.csv("csv data/rfbin_fw_vivi2.csv")
rfreg_df_vivi2 <- read.csv("csv data/rfreg_df_vivi2.csv")
rfbin_df_vivi2 <- read.csv("csv data/rfbin_df_vivi2.csv")
rfreg_mid_vivi2 <- read.csv("csv data/rfreg_mid_vivi2.csv")
rfbin_mid_vivi2 <- read.csv("csv data/rfbin_mid_vivi2.csv")
#
train_data_df2 <- read.csv("csv data/train_data_df2.csv")
test_data_df2 <- read.csv("csv data/test_data_df2.csv")
test_data_fw2 <- read.csv("csv data/test_data_fw2.csv") 
train_data_fw2 <- read.csv("csv data/train_data_fw2.csv")
train_data_mid2 <- read.csv("csv data/train_data_mid2.csv")
test_data_mid2 <- read.csv("csv data/test_data_mid2.csv")
train_data_tot2 <- read.csv("csv data/train_data_tot2.csv")
test_data_tot2 <- read.csv("csv data/test_data_tot2.csv")
#
df_data <- read.csv("csv data/df_data.csv")
mid_data <- read.csv("csv data/mid_data.csv")
fw_data <- read.csv("csv data/fw_data.csv")
tot_data <- read.csv("csv data/tot_data.csv")
#
rfbin_fw2 <- readRDS("models/rfbin_fw2.rds")
rfbin_df2 <- readRDS("models/rfbin_df2.rds")
rfbin_mid2 <- readRDS("models/rfbin_mid2.rds")
rfbin_tot2 <- readRDS("models/rfbin_tot2.rds")
rfreg_fw2 <- readRDS("models/rfreg_fw2.rds")
rfreg_df2 <- readRDS("models/rfreg_df2.rds")
rfreg_mid2 <- readRDS("models/rfreg_mid2.rds")
rfreg_tot2 <- readRDS("models/rfreg_tot2.rds")
# load these for the data from Transfermarkt and Fbref
all_val <- read.csv("csv data/all_val.csv")
final_combined_data <- read.csv("csv data/final_combined_data.csv")
joined_data = read.csv("csv data/team_data_pca.csv")
```

Define helper function for Variable Importance and vivi Matrix
```{r}
run_vivi_analysis <- function(rf_model, data, response) {
  vars <- head(sort(rf_model$variable.importance, decreasing = TRUE), 40)  # Top 40 important vars.
  
  vivi_matrix <- vivi(
    data = data,
    fit = rf_model,
    response = response,
    vars = names(vars)
  )
  
  return(vivi_matrix)
}
```

helper function numerische variable modell
```{r}
perform_grid_search_numeric <- function(mtry_values, num_trees_values, train_data, test_data, target_variable) {
  
  # Ensure target variable exists
  if (!(target_variable %in% colnames(train_data))) {
    stop(paste("Error: Target variable", target_variable, "not found in training data"))
  }
  
  if (!(target_variable %in% colnames(test_data))) {
    stop(paste("Error: Target variable", target_variable, "not found in test data"))
  }

  # Remove `gb_age` if it exists
  if ("gb_age" %in% colnames(train_data)) {
    train_data <- train_data %>% dplyr::select(-gb_age)
  }
  if ("gb_age" %in% colnames(test_data)) {
    test_data <- test_data %>% dplyr::select(-gb_age)
  }

  # Ensure both datasets have the same features
  common_columns <- intersect(colnames(train_data), colnames(test_data))
  train_data <- train_data[, common_columns]
  test_data <- test_data[, common_columns]

  # Move target variable to the first column
  train_data <- train_data %>% select(target_variable, everything())
  test_data <- test_data %>% select(target_variable, everything())

  # Convert target variable to numeric
  train_data[[target_variable]] <- as.numeric(train_data[[target_variable]])
  test_data[[target_variable]] <- as.numeric(test_data[[target_variable]])

  # Remove target variable from features before training
  train_features <- train_data %>% select(-target_variable)
  test_features <- test_data %>% select(-target_variable)

  # Initialize results dataframe
  tuning_results <- data.frame(
    mtry = integer(),
    num_trees = integer(),
    mse = numeric(),
    rmse = numeric(),
    mae = numeric(),
    r_squared = numeric()
  )

  # Perform grid search
  for (mtry in mtry_values) {
    for (num_trees in num_trees_values) {
      
      # Train Random Forest regression model
      rf_model <- ranger(
        formula = as.formula(paste(target_variable, "~ .")),
        data = train_data,
        num.trees = num_trees,
        mtry = mtry,
        importance = 'impurity',
        seed = 42
      )
      
      # Make predictions
      predictions <- predict(rf_model, data = test_features)$predictions
      actual <- test_data[[target_variable]]

      # Compute regression metrics
      mse_value <- mse(actual, predictions)
      rmse_value <- rmse(actual, predictions)
      mae_value <- mae(actual, predictions)
      r_squared_value <- cor(actual, predictions)^2  # R2 score
      
      # Store results
      new_row <- data.frame(
        mtry = mtry,
        num_trees = num_trees,
        mse = mse_value,
        rmse = rmse_value,
        mae = mae_value,
        r_squared = r_squared_value
      )
      
      tuning_results <- rbind(tuning_results, new_row)
    }
  }

  return(tuning_results)
}

```

helper function best model and performance with rf with binary model
```{r}
# Define hyperparameter grids
mtry_values <- c( 50, 70, 90, 100, 110, 120, 130, 151,179)#
num_trees_values <- c( 50, 100, 200, 300, 400, 500, 1000)


perform_grid_search <- function(mtry_values, num_trees_values, train_data, test_data) {
  
  # Drop Transfer_Value if present
  if ("Transfer_Value" %in% colnames(train_data)) {
    train_data <- train_data %>% select(-Transfer_Value)
  }
  if ("Transfer_Value" %in% colnames(test_data)) {
    test_data <- test_data %>% select(-Transfer_Value)
  }

  # Ensure that target variable is a factor
  train_data$gb_age <- as.factor(train_data$gb_age)
  test_data$gb_age <- as.factor(test_data$gb_age)

  # Prepare metrics data frame
  tuning_results <- data.frame(
    mtry = integer(),
    num_trees = integer(),
    accuracy = numeric(),
    auc = numeric(),
    predicted_0_actual_0 = integer(),
    predicted_1_actual_1 = integer(),
    predicted_0_actual_1 = integer(),
    predicted_1_actual_0 = integer()
  )

  # Define class and overall metrics manually
  class_metrics_names <- c(
    "Sensitivity", "Specificity", "Pos Pred Value", "Neg Pred Value",
    "Precision", "Recall", "F1", "Prevalence", "Detection Rate",
    "Detection Prevalence", "Balanced Accuracy"
  )
  overall_metrics_names <- c(
    "Accuracy", "Kappa", "AccuracyLower", "AccuracyUpper",
    "AccuracyNull", "AccuracyPValue", "McnemarPValue"
  )

  for (metric_name in c(class_metrics_names, overall_metrics_names)) {
    tuning_results[[metric_name]] <- numeric()
  }

  # Grid search loop for Random Forest hyperparameters
  for (mtry in mtry_values) {
    for (num_trees in num_trees_values) {

      #Train models
      rf_model <- ranger(
        formula = gb_age ~ .,
        data = train_data,
        num.trees = num_trees,
        mtry = mtry,
        importance = 'impurity',
        probability = TRUE,
        seed = 42
      )

      # Predict probabilities and classes
      predictions <- predict(rf_model, data = test_data)
      predicted_classes <- factor(ifelse(predictions$predictions[, "1"] >= 0.5, "1", "0"), levels = c("0", "1"))
      actual <- factor(test_data$gb_age, levels = c("0", "1"))

      # Confusion matrix 
      conf_matrix <- caret::confusionMatrix(predicted_classes, actual)

      # Do AUC
      positive_probs <- predictions$predictions[, "1"]
      binary_roc <- pROC::roc(response = actual, predictor = positive_probs, levels = c("0", "1"))
      auc_value <- pROC::auc(binary_roc)

      # Extract metrics
      overall_metrics <- conf_matrix$overall
      class_metrics <- conf_matrix$byClass
      confusion_counts <- as.vector(conf_matrix$table)

      new_row <- data.frame(
        mtry = mtry,
        num_trees = num_trees,
        accuracy = overall_metrics["Accuracy"],
        auc = auc_value,
        predicted_0_actual_0 = confusion_counts[1],
        predicted_1_actual_1 = confusion_counts[4],
        predicted_0_actual_1 = confusion_counts[3],
        predicted_1_actual_0 = confusion_counts[2]
      )

      # Add metrics
      for (metric_name in class_metrics_names) {
        new_row[[metric_name]] <- ifelse(metric_name %in% names(class_metrics), class_metrics[[metric_name]], NA)
      }
      for (metric_name in overall_metrics_names) {
        new_row[[metric_name]] <- ifelse(metric_name %in% names(overall_metrics), overall_metrics[[metric_name]], NA)
      }

      tuning_results <- rbind(tuning_results, new_row)
    }
  }

  return(tuning_results)
}
```

helper function for best hyperparameters
```{r}
get_best_hyperparameters <- function(df) {

  filtered_df <- df
  
  # Check if there are valid rows
  if (nrow(filtered_df) > 0) {
    if ("accuracy" %in% colnames(filtered_df)) {
      # select the row with the highest accuracy
      row_df <- filtered_df[which.max(filtered_df$accuracy), ]
    } else if ("rmse" %in% colnames(filtered_df)) {
      #  select the row with the lowest RMSE
      row_df <- filtered_df[which.min(filtered_df$rmse), ]
    } else {
      # Ifaccuracy and RMSE dont exist, return NA
      return(data.frame(Best_mtry = NA, Best_num_trees = NA))
    }
    
    # Extract optimal hyperparameters
    best_mtry <- row_df$mtry
    best_num_trees <- row_df$num_trees
  } else {
    # If no valid rows exist, NA
    best_mtry <- NA
    best_num_trees <- NA
  }
  
  return(data.frame(Best_mtry = best_mtry, Best_num_trees = best_num_trees))
}

```

helper function for training the model with hyperparam
```{r}
train_ranger_model_bin <- function(num_trees, mtry,data) {
  if (!is.na(num_trees) & !is.na(mtry)) {
    model <- ranger(
      formula = gb_age ~ ., 
      data = data, 
      num.trees = num_trees, 
      mtry = mtry, 
      importance = 'impurity', 
      probability = TRUE,
      seed = 42
    )
    return(model)
  } else {
    return(NULL)  
  }
}

train_ranger_model_reg <- function(num_trees, mtry,data) {
  if (!is.na(num_trees) & !is.na(mtry)) {
    model <- ranger(
      formula = Transfer_Value ~ ., 
      data = data, 
      num.trees = num_trees, 
      mtry = mtry, 
      importance = 'impurity', 
      probability = FALSE,
      seed = 42
    )
    return(model)
  } else {
    return(NULL)  
  }
}

```


no age and not market value and gb age in it do hyperparameter tuning and all
```{r}
# Perform grid search for classification 
rf_result_df_bin2  <- perform_grid_search(mtry_values, num_trees_values, train_data_df2, test_data_df2)
rf_result_mid_bin2 <- perform_grid_search(mtry_values, num_trees_values, train_data_mid2, test_data_mid2)
rf_result_fw_bin2  <- perform_grid_search(mtry_values, num_trees_values, train_data_fw2, test_data_fw2)
rf_result_tot_bin2 <- perform_grid_search(mtry_values, num_trees_values, train_data_tot2, test_data_tot2)

# Perform grid search for regression models 
rf_result_fw_reg2  <- perform_grid_search_numeric(mtry_values, num_trees_values, train_data_fw2, test_data_fw2, "Transfer_Value")
rf_result_mid_reg2 <- perform_grid_search_numeric(mtry_values, num_trees_values, train_data_mid2, test_data_mid2, "Transfer_Value")
rf_result_df_reg2  <- perform_grid_search_numeric(mtry_values, num_trees_values, train_data_df2, test_data_df2, "Transfer_Value")

rf_result_tot_reg2 <- perform_grid_search_numeric(
  mtry_values, num_trees_values,
  train_data_tot2[, names(train_data_tot2) != "gb_age"],
  test_data_tot2[, names(test_data_tot2) != "gb_age"],  
  "Transfer_Value"
)

fw_data2reg <- fw_data[, !(names(fw_data) %in% c("gb_age", "Age"))]
mid_data2reg <- mid_data[, !(names(mid_data) %in% c("gb_age", "Age"))]
df_data2reg <- df_data[, !(names(df_data) %in% c("gb_age", "Age"))]
tot_data2reg <- tot_data[, !(names(tot_data) %in% c("gb_age", "Age"))]

fw_data2bin <- fw_data[, !(names(fw_data) %in% c("Transfer_Value", "Age"))]
mid_data2bin <- mid_data[, !(names(mid_data) %in% c("Transfer_Value", "Age"))]
df_data2bin <- df_data[, !(names(df_data) %in% c("Transfer_Value", "Age"))]
tot_data2bin <- tot_data[, !(names(tot_data) %in% c("Transfer_Value", "Age"))]
# Need for vivd later, for models on all data

#####################
#hyperparameter tuning

best_df_reg2 <- get_best_hyperparameters(rf_result_df_reg2)
best_mid_reg2 <- get_best_hyperparameters(rf_result_mid_reg2)
best_fw_reg2 <- get_best_hyperparameters(rf_result_fw_reg2)
best_tot_reg2 <- get_best_hyperparameters(rf_result_tot_reg2)
best_df_bin2 <- get_best_hyperparameters(rf_result_df_bin2)
best_mid_bin2 <- get_best_hyperparameters(rf_result_mid_bin2)
best_fw_bin2 <- get_best_hyperparameters(rf_result_fw_bin2)
best_tot_bin2 <- get_best_hyperparameters(rf_result_tot_bin2)

#####################
#fit models after hyperparameter tuning

rfbin_fw2 <- train_ranger_model_bin(400, 110, fw_data2bin)
rfbin_df2 <- train_ranger_model_bin(300, 120, df_data2bin)
rfbin_mid2 <- train_ranger_model_bin(400, 100, mid_data2bin)
rfbin_tot2 <- train_ranger_model_bin(400, 70, tot_data2bin)

rfreg_fw2 <- train_ranger_model_reg(400, 90,fw_data2reg)
rfreg_df2 <- train_ranger_model_reg(100, 90,df_data2reg)
rfreg_mid2 <- train_ranger_model_reg(500, 130,mid_data2reg)
rfreg_tot2 <- train_ranger_model_reg(500, 70,tot_data2reg)
#########
#run vivi heatmaps

rfbin_fw_vivi2 <- run_vivi_analysis(rfbin_fw2, fw_data2bin, "gb_age")
rfbin_df_vivi2 <- run_vivi_analysis(rfbin_df2, df_data2bin, "gb_age")
rfbin_mid_vivi2 <- run_vivi_analysis(rfbin_mid2, mid_data2bin, "gb_age")
rfbin_tot_vivi2 <- run_vivi_analysis(rfbin_tot2, tot_data2bin, "gb_age")

# Regression models 
rfreg_fw_vivi2 <- run_vivi_analysis(rfreg_fw2, fw_data2reg, "Transfer_Value")
rfreg_df_vivi2 <- run_vivi_analysis(rfreg_df2, df_data2reg, "Transfer_Value")
rfreg_mid_vivi2 <- run_vivi_analysis(rfreg_mid2, mid_data2reg, "Transfer_Value")
rfreg_tot_vivi2 <- run_vivi_analysis(rfreg_tot2, tot_data2reg, "Transfer_Value")

###
```

helper functionvariable importance for both models bin/reg and the varimp function
```{r}
### Function to Compute and Combine Variable Importance for Four Models reg and bin
compute_var_importance <- function(model_fw, model_mid, model_df, model_tot, group_name) {

  varimp_fw <- as.data.frame(sort(model_fw$variable.importance, decreasing = TRUE))
  varimp_mid <- as.data.frame(sort(model_mid$variable.importance, decreasing = TRUE))
  varimp_df <- as.data.frame(sort(model_df$variable.importance, decreasing = TRUE))
  varimp_tot <- as.data.frame(sort(model_tot$variable.importance, decreasing = TRUE))

  colnames(varimp_fw) <- c("Importance_fw")
  colnames(varimp_mid) <- c("Importance_mid")
  colnames(varimp_df) <- c("Importance_df")
  colnames(varimp_tot) <- c("Importance_tot")

  # Create ranks for each model
  varimp_fw$Rank_fw <- rank(-varimp_fw$Importance_fw)
  varimp_mid$Rank_mid <- rank(-varimp_mid$Importance_mid)
  varimp_df$Rank_df <- rank(-varimp_df$Importance_df)
  varimp_tot$Rank_tot <- rank(-varimp_tot$Importance_tot)

  # Add variable names to each data frame
  varimp_fw$Variable <- rownames(varimp_fw)
  varimp_mid$Variable <- rownames(varimp_mid)
  varimp_df$Variable <- rownames(varimp_df)
  varimp_tot$Variable <- rownames(varimp_tot)

  # Join all four results together by the variable names
  combined_varimp <- full_join(varimp_fw, varimp_mid, by = "Variable") %>%
    full_join(varimp_df, by = "Variable") %>%
    full_join(varimp_tot, by = "Variable") %>%
    dplyr::mutate(
      Rank_Avg = (Rank_fw + Rank_mid + Rank_df + Rank_tot) / 4,  # Avrg rank across models
      Rank_Max = pmax(Rank_fw, Rank_mid, Rank_df, Rank_tot, na.rm = TRUE), 
      Rank_Min = pmin(Rank_fw, Rank_mid, Rank_df, Rank_tot, na.rm = TRUE)  
    )
  
  print(paste("Variable Importance for:", group_name))
  View(combined_varimp)
  
  return(combined_varimp)
}
```

variable imprtance table for binary classification of new models
```{r}
combined_varimp_bin2 <- compute_var_importance(rfbin_fw2, rfbin_mid2, rfbin_df2, rfbin_tot2,"bin names")

combined_varimp_reg2 <- compute_var_importance(rfreg_fw2, rfreg_mid2, rfreg_df2, rfreg_tot2,"reg names")

regbin = left_join(combined_varimp_reg2 %>%
  select(Variable,Rank_tot),combined_varimp_bin2 %>%
  select(Variable,Rank_tot), by = "Variable")
cor(regbin$Rank_tot.x, regbin$Rank_tot.y, use = "complete.obs")

```

heatmap comp_from_to 
Not int work, extra plot
```{r}
heatmap_data = pdp_comp %>%
  separate(Comp_from_to, into = c("Comp_from", "Comp_to"), sep = "_")

ggplot(heatmap_data, aes(x = Comp_from, y = Comp_to, fill = yhat)) +
  geom_tile() +
  geom_text(aes(label = round(yhat, 4)), color = "white", size = 4) +
  scale_fill_viridis_c(option = "magma") +  # Color scale
  labs(title = "Partial Dependence Heatmap: Comp vs Comp_going",
       x = "from",
       y = "to",
       fill = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Save all the important datasets as CSV files
```{r}
# write.csv(mid_data2reg, "csv data/mid_data2reg.csv", row.names = FALSE)
# write.csv(df_data2reg, "csv data/df_data2reg.csv", row.names = FALSE)
# write.csv(tot_data2reg, "csv data/vtot_data2reg.csv", row.names = FALSE)
# write.csv(fw_data2reg, "csv data/fw_data2reg.csv", row.names = FALSE)
# write.csv(fw_data2bin, "csv data/fw_data2bin.csv", row.names = FALSE)
# write.csv(mid_data2bin, "csv data/mid_data2bin.csv", row.names = FALSE)
# write.csv(df_data2bin, "csv data/df_data2bin.csv", row.names = FALSE)
# write.csv(tot_data2bin, "csv data/tot_data2bin.csv", row.names = FALSE)
# #
# write.csv(combined_varimp_bin2, "csv data/combined_varimp_bin2.csv", row.names = FALSE)
# write.csv(combined_varimp_reg2, "csv data/combined_varimp_reg2.csv", row.names = FALSE)
# write.csv(regbin, "csv data/regbin.csv", row.names = FALSE)
# #models
# saveRDS(rfbin_fw2, file = "models/rfbin_fw2.rds")
# saveRDS(rfbin_df2, file = "models/rfbin_df2.rds")
# saveRDS(rfbin_mid2, file = "models/rfbin_mid2.rds")
# saveRDS(rfbin_tot2, file = "models/rfbin_tot2.rds")
# saveRDS(rfreg_fw2, file = "models/rfreg_fw2.rds")
# saveRDS(rfreg_df2, file = "models/rfreg_df2.rds")
# saveRDS(rfreg_mid2, file = "models/rfreg_mid2.rds")
# saveRDS(rfreg_tot2, file = "models/rfreg_tot2.rds")
# #
# write.csv(rfreg_tot_vivi2, "csv data/rfreg_tot_vivi2.csv")
# write.csv(rfbin_tot_vivi2, "csv data/rfbin_tot_vivi2.csv")
# write.csv(rfreg_fw_vivi2, "csv data/rfreg_fw_vivi2.csv")
# write.csv(rfbin_fw_vivi2, "csv data/rfbin_fw_vivi2.csv")
# write.csv(rfreg_df_vivi2, "csv data/rfreg_df_vivi2.csv")
# write.csv(rfbin_df_vivi2, "csv data/rfbin_df_vivi2.csv")
# write.csv(rfreg_mid_vivi2, "csv data/rfreg_mid_vivi2.csv")
# write.csv(rfbin_mid_vivi2, "csv data/rfbin_mid_vivi2.csv")
```